Formula 1 2022 Regulation Changes

### 2024-04-22 paper/paper.qmd
made some updates to results and discussion section

---
title: "Formula 1 2022 Regulation Changes"
subtitle: "How did the 2022 F1 Regulations change the sport for better and for worse"
author: 
  - Michael Fang
thanks: "Code and data are available at: https://github.com/fanger2791/F1_Regulations."
date: today
date-format: long
abstract: "As Formula 1's new era of hybrid cars develops, a fresh range of rules and regulations have been put in place for the teams to follow attempting to create a closer, more competitive championship. This paper will investigate the impact of the 2022 Formula 1 regulations, designed to intensify the competitive landscape and elevate the racing experience through technical and fiscal adjustments. It examines lap times and overtaking dynamics, while the analysis zeroes in on the performance trajectories of the sport’s leading contenders: Max Verstappen, Lewis Hamilton, and Charles Leclerc across varied circuit types from 2021 to 2022. It focuses on whether these regulatory modifications have indeed leveled the playing field, creating tighter competition and mitigating the performance differences between teams or have increased the gap even more. This research not only examines the immediate effects of the 2022 changes but also considers the balance between innovation and competition in Formula 1, thereby contributing to the ongoing dialogue on how to shape the future of this high-stakes sport."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

variable_data <- data.frame( 
  Circuit_Name = c("Saudi Arabia", 
               "Netherlands"), 
  Track_Type = c("Top Speed",
                  "High Downforce"), 
  Tyre_Wear = c("High", 
              "Low")
)
variable_data |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

@@

@@

>>>>

---
title: "Formula 1 2022 Regulation Changes"
subtitle: "How did the 2022 F1 Regulations change the sport"
author: 
  - Michael Fang
thanks: "Code and data are available at: https://github.com/fanger2791/F1_Regulations."
date: today
date-format: long
abstract: "As Formula 1's new era of hybrid cars develops, a fresh range of rules and regulations have been put in place for the teams to follow attempting to create a closer, more competitive championship. This paper will investigate the impact of the 2022 Formula 1 regulations, designed to intensify the competitive landscape and elevate the racing experience through technical and fiscal adjustments. It examines lap times and overtaking dynamics, while the analysis zeroes in on the performance trajectories of the sport’s leading contenders: Max Verstappen, Lewis Hamilton, and Charles Leclerc across varied circuit types from 2021 to 2022. It focuses on whether these regulatory modifications have indeed leveled the playing field, creating tighter competition and mitigating the performance differences between teams or have increased the gap even more. This research not only examines the immediate effects of the 2022 changes but also considers the balance between innovation and competition in Formula 1, thereby contributing to the ongoing dialogue on how to shape the future of this high-stakes sport."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

variable_data <- data.frame( 
  CircuitName = c("Saudi Arabia", 
               "Netherlands"), 
  TrackType = c("Top Speed",
                  "High Downforce"), 
  TyreWear = c("High", 
              "Low")
)
variable_data |>
 slice(1:10) |>
  kable(digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

@@

## Fastest Qualifying Lap Comparison Between 2021 and 2022 in the Netherlands - Max Verstappen, Lewis Hamilton and Charles Leclerc

@fig-verspeed-netherlands, @fig-hamspeed-netherlands, @fig-lecspeed-netherlands illustrate a comparison of Max Verstappen, Lewis Hamilton and Charles Leclerc's speeds during qualifying sessions in the Netherlands for the 2021 and 2022 Formula 1 seasons, giving insight into the performance changes due to new regulations.

Like the previous graphs, the top graph shows a general comparison of speeds throughout a lap, marked by time in seconds. The lines trace the speed of the car at different points of the lap, with two different lines, one representing 2021 and the other for 2022. We observe that the speed profiles for both years are quite similar in their fluctuations, which correspond to the acceleration and deceleration points around the lap. However, the line for 2022 occasionally edges above the line from 2021, especially in the second half of the graph. This indicates that driver's car in 2022 was able to achieve higher speeds at certain points during the lap, potentially benefiting from the reduced aerodynamic drag on the straights due to the new regulations.

The bottom graph provides a focused comparison of speeds at a critical section of the track involving turns 11 and 12. The 2022 speeds dip below those of 2021 as the car approaches and navigates through the corners. This supports the notion that while the 2022 F1 cars may gain speed on the straights, they are likely to lose some of that advantage in slower, high-downforce corners. The increased weight of the 2022 cars, coupled with the change to a ground-effect aerodynamic model that produces less downforce at low speeds, could result in slower cornering speeds. This slower speed through corners can be critical in sectors of the track where maintaining momentum is key, reflecting the trade-offs introduced by the new regulations.

Overall, these graphs suggest that the 2022 F1 cars could exhibit a complex performance profile that is a direct consequence of the trade-offs between straight-line speed and cornering agility introduced by the regulatory changes. The driver's telemetry data provides a practical example of these effects, showing possible gains on the straights and challenges in the corners.

@@

## Potential Issues with New Regulations

The 2022 Formula 1 regulations, while innovative and well-intentioned in their approach to improving race competitiveness and sustainability, manifest certain weaknesses that could undermine their effectiveness. A primary concern is the reliance on relatively short-term data to assess the impact of these regulations. This limited timeframe may not adequately reflect the full range of effects, particularly in how teams adapt strategically over longer periods. Such an evaluation might overlook emerging trends or delayed consequences, leading to potentially skewed interpretations of the regulations' success or failure.

Another notable issue is the increased complexity and cost associated with the new technical requirements, despite the implementation of cost caps. These regulations demand significant changes to car designs, particularly in aerodynamics and tire usage, which can disproportionately affect smaller teams with fewer resources for rapid adaptation and innovation. This situation can paradoxically widen the competitive gap the regulations aim to narrow.

Moreover, while the switch to ground effect aerodynamics aims to reduce the 'dirty air' problem, it simultaneously introduces new challenges in handling and tire degradation, potentially leading to unanticipated competitive disparities and affecting the overall quality of racing.

## How should F1 Proceed in the Future

As Formula 1 continues to evolve under the 2022 regulations, several areas require further exploration to fully understand and optimize the impacts of these changes. First and foremost, a longer-term assessment is crucial. The initial seasons under new regulations provide valuable data, but the true test will come as teams fully adapt and mature within the new framework. This will allow for a more accurate evaluation of the regulations' long-term sustainability and their effect on the competitive balance. Investigating over multiple seasons will help determine if smaller teams can indeed close the gap with top-tier teams or if financial and technical disparities continue to dominate outcomes.

Additionally, further research into the environmental impacts of the sport is needed, particularly concerning the real-world benefits of using sustainable fuels and advanced materials. As the global emphasis on reducing carbon footprints intensifies, F1 could serve as a pioneering platform for innovative green technologies. This includes not only the fuels used but also the materials and methods employed in manufacturing and transporting race equipment.

Technological development should also be a focal point. The potential of emerging technologies such as artificial intelligence and advanced simulation tools to enhance car performance and race strategies should be explored. These technologies could help teams optimize their designs more efficiently and adapt to regulations more effectively.

Furthermore, enhancing fan engagement through adjustments in race formats, accessibility of content, and interactive platforms will be vital. As the sport aims to attract a broader, more diverse audience, understanding the preferences and consumption patterns of this audience will be essential.

# Conclusion
